---
title: "Assignment_03_27"
author: "Nilesh Dighe"
date: "March 27, 2015"
output: html_document
---


```{r echo=FALSE}
setwd("C:/Users/NNDIGH/Desktop/coursera/Machine Learning/Home Work/Machine Learning_HW")
library(ggplot2)
library(caret)
library(randomForest)
trainset<- read.csv("pml-training.csv", na.strings=c("NA", "#DIV/0!"))
testset<- read.csv("pml-testing.csv", na.strings=c("NA", "#DIV/0!"))
dim(trainset)
dim(testset)
```

*Removing columns with "NA" values*
```{r echo=FALSE}
train1<- trainset[,colSums(is.na(trainset)) <2]
dim(train1)
```

*remove columns 1 & 4 from train1 data set*
```{r}
train2<- train1[,-1]
train3<- train2[,-4]
str(train3)
dim(train3)
```
*change trainControl function as listed below*
```{r}
ctrl1 <- trainControl(method = "crossvalidate", number=3, repeats = 3)
```
*split data into training and testing sets*
```{r}
inTrain3<- createDataPartition(y=train3$classe, p=0.75, list=FALSE)
training3<- train3[inTrain3,]
testing3<- train3[-inTrain3,]
dim(training3); dim(testing3)
```
*fine tune model by including top 20 variables obtained via varImp function above*
```{r}
modelFit5<- train(classe~ raw_timestamp_part_1 + num_window + roll_belt + pitch_forearm + magnet_dumbbell_z
                         + yaw_belt + magnet_dumbbell_y + pitch_belt + roll_forearm + accel_dumbbell_y
                         + roll_dumbbell + accel_belt_z + accel_forearm_x + magnet_dumbbell_x + magnet_belt_y
                         + total_accel_dumbbell + magnet_belt_z + accel_dumbbell_z + magnet_forearm_z + magnet_arm_x, method="rf", data=training3, trainControl=ctrl1)
modelFit5
```
*run cross validation and print data in table*
```{r}
pred<- predict(modelFit5,testing3)
table(pred,testing3$classe)
```


*run confusion matrix function*
```{r}
confusionMatrix(pred, testing3$classe)
```

*make prediction in the testset and place results in table and vectors*
```{r}
predict_test<- predict(modelFit5,testset)
predict_test[1:20]
table(predict_test)
```