---
title: "Prediction Model"
author: "Nilesh Dighe"
date: "April 25, 2015"
output: html_document
---

##Steps taken in building the _classe_ prediction model
  + Data clean up
  + Create training and testing sets
  + Run full model
  + Identify most important variables
  + Check for normality of the variables
  + Fine tune model by only including the important variables
  + Test model on validation or new testing set
  + Estimate Accuracy and out of sample error
  + Make prediction on the test set and submit answers as part of the Coursera submission project


```{r include=FALSE}
library(ggplot2)
library(caret)
library(randomForest)
library(dplyr)
trainset<- read.csv("pml-training.csv", na.strings=c("NA", "#DIV/0!"))
testset<- read.csv("pml-testing.csv", na.strings=c("NA", "#DIV/0!"))
str(trainset)
head(trainset)
dim(trainset)
train1<- trainset[,colSums(is.na(trainset)) <1]
dim(train1)
str(train1)
train2<- train1[,-1]
train3<- train2[,-4]
dim(train3)
```

*Partition training data set into train and test sets*
```{r}
inTrain3<- createDataPartition(y=train3$classe, p=0.75, list=FALSE)
training3<- train3[inTrain3,]
testing3<- train3[-inTrain3,]
dim(training3); dim(testing3)
```

*Histograms & qqplots*

```{r}
par(mfrow=c(1,2));hist(training3$raw_timestamp_part_1); qqnorm(training3$raw_timestamp_part_1)
par(mfrow=c(1,2));hist(training3$num_window); qqnorm(training3$num_window)
```



*Change trainControl function as listed below*
```{r}
ctrl1 <- trainControl(method = "crossvalidate", number=3, repeats = 3)
```


###Run complete model and identiy top 20 most importnat variables##
###Full model: modelFit3<- train(classe~ ., method="rf", data=training3, trainControl=ctrl1)
###Identify top 20 variables using varImp(modelFit3)
**Variable**        |**Overall**
--------------------|------
raw_timestamp_part_1|100.000
num_window          | 55.406
roll_belt           | 46.066
pitch_forearm       | 28.259
magnet_dumbbell_z   | 21.192
magnet_dumbbell_y   | 18.707
yaw_belt            | 17.221
pitch_belt          | 15.997
roll_forearm        | 14.042
accel_dumbbell_y    |  8.576
roll_dumbbell       |  7.931
magnet_dumbbell_x   |  6.868
accel_belt_z        |  6.413
total_accel_dumbbell|  6.304
magnet_belt_y       |  5.781
accel_forearm_x     |  5.655
magnet_belt_z       |  5.464
accel_dumbbell_z    |  5.103
yaw_dumbbell        |  3.142
gyros_belt_z        |  3.095

*Evaluate variables for trends using featureplot*
```{r}
featurePlot(x=training3[c("raw_timestamp_part_1","num_window","roll_belt","pitch_forearm")], y=training3$classe, plot="pairs")
```


###Fine tune model by including top 2 variables obtained via varImp function above###

```{r, warnings=FALSE, message=FALSE}
modelFit3<- train(classe~ raw_timestamp_part_1 + num_window, method="rf", data=training3, trainControl=ctrl1)
modelFit3
```


*Run cross validation and print data in table*
```{r, echo=FALSE}
pred<- predict(modelFit3,testing3)
table(pred,testing3$classe)
```


*Run confusion matrix function to estimate accuracy & out of sample error*
```{r,echo=FALSE}
cm<- confusionMatrix(pred, testing3$classe)
cm
AccuracyStats<- data.frame(cm$overall)
OutofSampleError<- 1-AccuracyStats[1,1]
OutofSampleError
e <- OutofSampleError * 100
paste0("Out of sample error estimation: ", round(e, digits = 2), "%")
```
#Make prediction in the testset and place results in table and vectors
```{r}
predict_test<- predict(modelFit3,testset)
predict_test[1:20]
table(predict_test)
```
